---
title: "Choosing an Execution Mode"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing an Execution Mode}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`rfastloess` provides three execution modes for different data sizes and processing needs:

![Execution mode comparison](../man/figures/adapter_comparison.svg)

| Mode | Class | Use Case | Memory | Key Features |
| --- | --- | --- | --- | --- |
| **Batch** | `Loess` | Complete datasets in memory | Entire dataset | All features: CI, PI, CV |
| **Streaming** | `StreamingLoess` | Large files (>100K points) | One chunk at a time | Chunked processing with overlap |
| **Online** | `OnlineLoess` | Real-time / live data | Fixed sliding window | Point-by-point incremental updates |

## Quick Decision Guide

- **Data fits in memory and you need intervals or CV** → [Batch Mode](batch.html)
- **Data is too large for memory or arrives in file chunks** → [Streaming Mode](streaming.html)
- **Data arrives point-by-point in real time** → [Online Mode](online.html)

```{r session-info}
sessionInfo()
```
