---
title: "Choosing an Execution Mode"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing an Execution Mode}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`rfastloess` provides three execution modes for different data sizes and
processing needs:

![Execution mode comparison](../man/figures/adapter_comparison.svg)

| Mode | Class | Use Case | Memory | Key Features |
| --- | --- | --- | --- | --- |
| **Batch** | `Loess` | Full dataset | Entire dataset | CI, PI, CV |
| **Streaming** | `StreamingLoess` | >100K points | Per chunk | With overlap |
| **Online** | `OnlineLoess` | Live / real-time | Sliding window | Per-point |

## Quick Decision Guide

| Situation | Mode |
| --- | --- |
| Data fits in memory; needs intervals or CV | Batch |
| Data too large for memory or arrives in chunks | Streaming |
| Data arrives point-by-point in real time | Online |

```{r session-info}
sessionInfo()
```
