---
title: "Alternative Software"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alternative Software}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

`rfastloess` and base R's `stats::loess()` both fit locally weighted
polynomials. Their defaults and result objects differ, so a useful numerical
comparison needs to match the fit settings first. This page shows how to do
that and where the two APIs offer different capabilities.

## Comparing fitted values

For a non-robust, one-dimensional local linear fit, use the same span and
polynomial degree, disable boundary padding, and request the direct surface.
`stats::loess()` uses the Gaussian family by default; `iterations = 0L`
disables robust reweighting in `Loess()`.

```{r alternative_software_gaussian}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 60)
y <- sin(x) + rnorm(length(x), sd = 0.2)

reference <- stats::loess(
    y ~ x, span = 0.5, degree = 1, family = "gaussian",
    control = stats::loess.control(surface = "direct")
)
result <- fit(
    Loess(
        fraction = 0.5, iterations = 0L, surface_mode = "direct",
        boundary_policy = "noboundary", parallel = FALSE
    ),
    x, y
)
max_difference <- max(abs(result$y - reference$fitted))
cat("Within 1e-10:", max_difference < 1e-10, "\n")
stopifnot(max_difference < 1e-10)
```

For robust fitting, R uses `family = "symmetric"`. Match its iteration
count and uncentered median absolute residual (`"mar"`) scale. The
zero-weight fallback is set explicitly for degenerate neighborhoods; this
example does not exercise that fallback:

```{r alternative_software_robust}
reference_robust <- stats::loess(
    y ~ x, span = 0.5, degree = 1, family = "symmetric",
    control = stats::loess.control(surface = "direct", iterations = 3)
)
result_robust <- fit(
    Loess(
        fraction = 0.5, iterations = 3L, surface_mode = "direct",
        boundary_policy = "noboundary", scaling_method = "mar",
        zero_weight_fallback = "return_original", parallel = FALSE
    ),
    x, y
)
robust_difference <- max(abs(result_robust$y - reference_robust$fitted))
cat("Within 1e-10:", robust_difference < 1e-10, "\n")
stopifnot(robust_difference < 1e-10)
```

These checks demonstrate agreement for the stated data and settings, not a
universal tolerance guarantee across all inputs. `stats::loess()$fitted` and
`fit(Loess(), x, y)$y` both follow the input observation order by default;
`outputs = "sorted"` is available when sorted LOESS output is preferred.

## Why defaults differ

| Setting | `stats::loess()` | `Loess()` |
| --- | --- | --- |
| Local neighborhood | `span = 0.75` | `fraction = 0.67` |
| Polynomial | `degree = 2` | `degree = "linear"` |
| Robust fitting | `family = "gaussian"` (off) | `iterations = 3L` |
| Boundary handling | No added padding | `boundary_policy = "extend"` |
| Residual scale for robust fits | Uncentered absolute residual (`"mar"`) | `scaling_method = "mad"` |
| Zero-weight fallback | No configurable option | `"use_local_mean"` |

`surface_mode = "direct"` and R's `surface = "direct"` remove interpolation
from the comparison. R's default is interpolation, as is the LOESS binding's.
The alternative settings above are for reference matching; the package
retains its own defaults for ordinary use. See `vignette("boundary")` and
`vignette("scaling")` for their effects.

## What this package adds

Compared with `stats::loess()`, `rfastloess` provides:

| Feature | `rfastloess` | `stats::loess()` |
| --- | :---: | :---: |
| Polynomial degree | 0-4 | 0-2 |
| Kernel functions | 7 options | Tricube only |
| Robustness weighting | 3 methods | Bisquare only |
| Scale estimation | MAD, MAR, mean | MAR for robust fits |
| Distance metrics | 6 options | Euclidean, optionally normalized |
| Boundary padding | 4 policies | None |
| Zero-weight fallback | 3 options | Not configurable |
| Automatic convergence | Yes | No |
| Built-in confidence / prediction interval bounds | Yes | SE available; bounds require calculation |
| Cross-validation for `fraction` | K-fold, LOOCV | No automatic span selection |
| Streaming / online modes | Yes | No |
| Parallel execution | Yes | No built-in parallel fitting |

Both support multivariate predictors, per-observation weights, prediction,
and prediction standard errors. See `vignette("concepts")` for an overview of
the additional options, or the [Benchmarks](benchmarks.html) article for
performance comparisons.
