---
title: "Batch Mode (Loess)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch Mode (Loess)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

> **See also:** [Choosing an Execution Mode](adapter-choice.html) for a comparison of all three modes.

The default mode. Processes the entire dataset at once and supports all features.

## When to Use

- Dataset fits comfortably in memory
- Need confidence/prediction intervals
- Need cross-validation
- One-shot analysis

## Parameters

| Parameter | Default | Description |
| --- | --- | --- |
| `fraction` | 0.67 | Neighbourhood size |
| `iterations` | 0 | Robustness iterations |
| `confidence_intervals` | `NULL` | CI coverage (e.g. 0.95) |
| `prediction_intervals` | `NULL` | PI coverage (e.g. 0.95) |
| `parallel` | `FALSE` | Enable CPU parallelism |

## Example

```{r}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(
    fraction = 0.5,
    iterations = 3,
    confidence_intervals = 0.95,
    prediction_intervals = 0.95,
    return_diagnostics = TRUE
)
result <- fit(model, x, y)

print(result$confidence_lower)
print(result$confidence_upper)
print(result$diagnostics$r_squared)
```

![Gap handling](../man/figures/gap_handling.svg)
