---
title: "Custom Weights"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Weights}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## How Custom Weights Work

Per-observation weights encode data quality directly into the LOESS fit. The effective weight of observation $j$ in a local fit centred at $x_i$ is:

$$w_{ij} = \text{custom\_weights}[j] \times K\!\left(\frac{d_{ij}}{h_i}\right) \times r_j$$

> **Note:** `custom_weights` applies in **Batch** mode only.

---

## When to Use Custom Weights

| Situation | Recommended weight |
| --- | --- |
| Point known to be erroneous | `0.0` — fully excluded |
| Unreliable sensor / low precision | `0.1 – 0.5` |
| Standard observation | `1.0` (default) |
| Carefully calibrated measurement | `> 1.0` |
| Measurement uncertainty $\sigma_i$ | $1 / \sigma_i^2$ |

---

## Suppress a Known Outlier

```{r custom_weights_1}
library(rfastloess)

x <- 1:10
y <- x * 2.0
y[6] <- 100.0              # spike at index 6

weights <- rep(1.0, 10)
weights[6] <- 0.0          # exclude the spike

model <- Loess(fraction = 0.5, iterations = 0L)
result <- fit(model, x, y, custom_weights = weights)
```

---

## Inverse-Variance Weights

```{r custom_weights_2}
library(rfastloess)
set.seed(42)
x <- seq(0, 10, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.5)

sigma <- 0.1 + 0.5 * abs(sin(x))
weights <- 1 / sigma^2

model <- Loess(fraction = 0.3)
result <- fit(model, x, y, custom_weights = weights)

plot(x, y, pch = 16, cex = 0.5 + weights / max(weights),
     col = "gray", main = "Inverse-Variance Weighted LOESS")
lines(result$x, result$y, col = "blue", lwd = 2)
```

---

## Combining with Robustness

```{r custom_weights_3}
library(rfastloess)
set.seed(42)
x <- 1:100
y <- sin(x / 10) + rnorm(100, sd = 0.3)

# Known bad region: indices 40–50
weights <- rep(1.0, 100)
weights[40:50] <- 0.1

# Also use robustness for unknown outliers
model <- Loess(fraction = 0.3, iterations = 3)
result <- fit(model, x, y, custom_weights = weights)
```

```{r session-info}
sessionInfo()
```
