---
title: "Polynomial Degree"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Polynomial Degree}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

At each target point, LOESS fits a polynomial to the neighbouring data using
weighted least squares. The `degree` parameter controls the order of that
polynomial.

![Degree comparison](../man/figures/degree_comparison.svg)

| Degree | Local Fit | Captures | Risk |
| --- | --- | --- | --- |
| `0` / `"constant"` | Weighted mean | Level only | Over-smooth; edge bias |
| `1` / `"linear"` | Weighted line | Trend (default) | Rarely overfits |
| `2` / `"quadratic"` | Parabola | Curvature | Overfits with small `fraction` |
| `3` / `"cubic"` | Cubic curve | Inflections | Requires larger `fraction` |
| `4` / `"quartic"` | Quartic curve | Fine structure | High variance, rare |

Degree can be specified as an integer (`0L`–`4L`) or as a string (`"constant"`,
`"linear"`, `"quadratic"`, `"cubic"`, `"quartic"`).

---

## Degree 0 — Local Constant

The fit at each point is a weighted mean. Produces very smooth results but
ignores local slope, introducing bias wherever the true function changes.

**Use when**: Maximum smoothness matters more than accuracy; computationally
cheapest.

```{r degree_1}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(degree = "constant", fraction = 0.5)
result <- fit(model, x, y)
cat("First 6 smoothed values (constant/Nadaraya-Watson):\n")
print(head(result$y))
```

---

## Degree 1 — Local Linear (Default)

Fits a weighted line through the neighbourhood. Removes first-order bias and
handles boundary regions correctly. The right choice for the vast majority of
applications.

**Use when**: Default; monotone or gently curved data; boundary accuracy
matters.

```{r degree_2}
model <- Loess(degree = "linear", fraction = 0.5)
result <- fit(model, x, y)
cat("First 6 smoothed values (linear local regression):\n")
print(head(result$y))
```

---

## Degree 2 — Local Quadratic

Fits a weighted parabola. Captures curvature in the data and reduces bias in
regions with strong local curvature, at the cost of higher variance.

**Use when**: Data has visible curvature or local extrema; use a larger
`fraction` to stabilise the fit.

```{r degree_3}
model <- Loess(degree = "quadratic", fraction = 0.5)
result <- fit(model, x, y)
cat("First 6 smoothed values (quadratic local regression):\n")
print(head(result$y))
```

---

## Degree 3 — Local Cubic

Fits a weighted cubic polynomial. Captures inflections and fine structure but
requires a sufficiently large neighbourhood to remain stable.

**Use when**: Data has clear S-shaped curves or multiple inflection points;
use `fraction` >= 0.5.

```{r degree_4}
model <- Loess(degree = "cubic", fraction = 0.6)
result <- fit(model, x, y)
cat("First 6 smoothed values (cubic local regression):\n")
print(head(result$y))
```

---

## Degree 4 — Local Quartic

Fits a weighted quartic polynomial. Rarely needed in practice; only useful for
capturing highly oscillatory local structure. Very prone to overfitting.

**Use when**: Fine oscillatory structure is physically meaningful and the
dataset is large; always cross-validate.

```{r degree_5}
model <- Loess(degree = "quartic", fraction = 0.7)
result <- fit(model, x, y)
cat("First 6 smoothed values (quartic local regression):\n")
print(head(result$y))
```

---

## Degree and Boundary Fallback

At the dataset boundary, the neighbourhood is one-sided and may not contain
enough points to support the requested degree. Setting `boundary_degree_fallback
= TRUE` automatically reduces the degree at boundaries to avoid instability.

```{r degree_6}
model <- Loess(degree = 2L, boundary_degree_fallback = TRUE)
result <- fit(model, x, y)
cat("First 6 smoothed values (quadratic, boundary fallback):\n")
print(head(result$y))
```

---

## Surface Mode

The `surface_mode` parameter controls whether LOESS evaluates the local
polynomial at every query point or at a sparser grid of vertices with Hermite
cubic interpolation in between.

| Mode | Behaviour | Speed | Accuracy |
| --- | --- | --- | --- |
| `"interpolation"` (default) | Evaluate at anchor vertices, blend via Hermite cubic | Faster | Slight approximation |
| `"direct"` | Evaluate at every query point | Exact | Full precision |

See [Multivariate LOESS](dimensions.html) for runnable `surface_mode` examples.

---

## Choosing a Degree

| Situation | Recommended Degree |
| --- | --- |
| Monotone trend, general purpose | `1` (default) |
| Maximum smoothness, speed | `0` |
| Clear peaks / valleys / inflections | `2` (with `fraction` >= 0.4) |
| S-shaped curves, multiple inflections | `3` (with `fraction` >= 0.5) |
| Fine oscillatory structure (rare) | `4` (with `fraction` >= 0.6, cross-validate) |
| Boundary accuracy is critical | `1` or `2` (not `0`) |
| Very small dataset (n < 50) | `1` |

> **Rule of thumb:** Start with `degree = 1` (default). Move to `degree = 2`
> only if you see systematic bias in regions of high curvature, and increase
> `fraction` at the same time.

```{r session-info}
sessionInfo()
```
