---
title: "Polynomial Degree"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Polynomial Degree}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Overview

At each target point, LOESS fits a polynomial to the neighbouring data using weighted least squares. The `degree` parameter controls the order of that polynomial.

![Degree comparison](../man/figures/degree_comparison.svg)

| Degree | Local Fit | Captures | Risk |
| --- | --- | --- | --- |
| `0` / `"constant"` | Weighted mean | Level only | Over-smooth, biased at edges |
| `1` / `"linear"` | Weighted line | Trend (default) | Rarely overfits |
| `2` / `"quadratic"` | Parabola | Curvature | Overfits with small `fraction` |
| `3` / `"cubic"` | Cubic curve | Inflections | Requires larger `fraction` |
| `4` / `"quartic"` | Quartic curve | Fine structure | High variance, rarely needed |

Degree can be specified as an integer (`0L`–`4L`) or as a string (`"constant"`, `"linear"`, `"quadratic"`, `"cubic"`, `"quartic"`).

---

## Degree 0 — Local Constant

The fit at each point is a weighted mean. Produces very smooth results but ignores local slope, introducing bias wherever the true function changes.

**Use when**: Maximum smoothness matters more than accuracy; computationally cheapest.

```{r degree_1}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(degree = "constant", fraction = 0.5)
result <- fit(model, x, y)
```

---

## Degree 1 — Local Linear (Default)

Fits a weighted line through the neighbourhood. Removes first-order bias and handles boundary regions correctly. The right choice for the vast majority of applications.

**Use when**: Default; monotone or gently curved data; boundary accuracy matters.

```{r degree_2}
model <- Loess(degree = "linear", fraction = 0.5)
result <- fit(model, x, y)
```

---

## Degree 2 — Local Quadratic

Fits a weighted parabola. Captures curvature in the data and reduces bias in regions with strong local curvature, at the cost of higher variance.

**Use when**: Data has visible curvature or local extrema; use a larger `fraction` to stabilise the fit.

```{r degree_3}
model <- Loess(degree = "quadratic", fraction = 0.5)
result <- fit(model, x, y)
```

---

## Degree 3 and 4 — Higher-Order Fits

Cubic and quartic local fits capture inflections and fine structure but require a sufficiently large neighbourhood to remain stable. Rarely needed in practice.

**Use when**: Highly non-linear local structure; combine with a larger `fraction`.

```{r degree_4}
model <- Loess(degree = "cubic", fraction = 0.6)
result <- fit(model, x, y)
```

---

## Degree and Boundary Fallback

At the dataset boundary, the neighbourhood is one-sided and may not contain enough points to support the requested degree. Setting `boundary_degree_fallback = TRUE` automatically reduces the degree at boundaries to avoid instability.

```{r degree_5}
model <- Loess(degree = 2L, boundary_degree_fallback = TRUE)
result <- fit(model, x, y)
```

---

## Choosing a Degree

| Situation | Recommended Degree |
| --- | --- |
| Flat or slowly varying signal | `0` |
| General purpose | `1` (default) |
| Visibly curved signal | `2` |
| Strong non-linearity, large dataset | `3` |
| Benchmark / exploratory only | `4` |

> **Rule of thumb:** Start with `degree = 1` (default). Move to `degree = 2` only if you see systematic bias in regions of high curvature, and increase `fraction` at the same time.
