---
title: "Multivariate LOESS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multivariate LOESS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Overview

Standard LOESS operates on a single predictor `x`. LOESS extends this to an **n-dimensional predictor space** by passing a matrix as `x`, enabling surface smoothing over spatial grids, time–altitude combinations, and similar multi-predictor datasets.

| Dimensions | Use Case | `x` Input Shape |
| --- | --- | --- |
| `1` | Time series, 1D signal (default) | Numeric vector |
| `2` | Spatial surface, 2-predictor model | `n × 2` matrix |
| `3+` | High-dimensional regression | `n × d` matrix |

> **Note:** Neighbourhood search scales with the number of dimensions. For `d ≥ 3` keep `fraction` small and consider the `"interpolation"` surface mode (the default) to reduce computation.

---

## 1D — Standard (Default)

Single predictor. No configuration required.

```{r}
library(rfastloess)
set.seed(42)
x <- seq(0, 10, length.out = 200)
y <- sin(x) + rnorm(200, sd = 0.2)

model <- Loess(fraction = 0.3)
result <- fit(model, x, y)
```

---

## 2D — Spatial Surface

Two predictors (e.g., latitude/longitude, time/altitude). Pass an `n × 2` matrix as `x`.

```{r}
set.seed(42)
n <- 100
lat <- runif(n, -pi, pi)
lon <- runif(n, -pi, pi)
z   <- sin(lat) + cos(lon) + rnorm(n, sd = 0.1)

x2d <- cbind(lat, lon)   # n × 2 matrix

model <- Loess(fraction = 0.3)
result <- fit(model, x2d, z)
```

---

## 3D and Higher

Pass an `n × d` matrix as `x`. Higher dimensions require a larger `fraction` to keep the local neighbourhood populated.

```{r}
set.seed(42)
n <- 200
x1 <- runif(n)
x2 <- runif(n)
x3 <- runif(n)
y  <- x1 + sin(2 * pi * x2) + x3^2 + rnorm(n, sd = 0.1)

x3d <- cbind(x1, x2, x3)   # n × 3 matrix

model <- Loess(fraction = 0.5)
result <- fit(model, x3d, y)
```

---

## Distance Metrics

In multivariate mode, neighbour distances can be computed with different metrics.

| Metric | Description | When to Use |
| --- | --- | --- |
| `"normalized"` | Each dimension scaled by its range (default) | Mixed-scale predictors |
| `"euclidean"` | Raw Euclidean distance | Same-scale predictors |
| `"weighted"` | Per-dimension scaling via `weighted_metric_weights` | Custom anisotropic distance |

```{r}
# Upweight time dimension relative to spatial dimension
x2d <- cbind(time_col, space_col)

model <- Loess(
    fraction = 0.3,
    distance_metric = "weighted",
    weighted_metric_weights = c(2.0, 1.0)  # time counts twice
)
result <- fit(model, x2d, y)
```

---

## Surface Mode

| Mode | Behaviour | Speed |
| --- | --- | --- |
| `"interpolation"` | Fit at grid vertices, interpolate elsewhere (default) | Fast |
| `"direct"` | Fit directly at every query point | Exact, slower |

```{r}
# Exact fit at every point — slower but no interpolation artefacts
model <- Loess(surface_mode = "direct", fraction = 0.3)
result <- fit(model, x2d, z)
```

For large 2D or 3D datasets use `"interpolation"` (default) and tune `cell` and `interpolation_vertices` to control grid resolution.

```{r}
model <- Loess(
    surface_mode = "interpolation",
    cell = 0.2,
    interpolation_vertices = 200L,
    fraction = 0.3
)
result <- fit(model, x2d, z)
```
