---
title: "GPU Backend"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GPU Backend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Overview

The batch `Loess` type can execute on a GPU-accelerated backend powered by `wgpu` (Vulkan / Metal / DX12).

> **Batch only.** GPU support applies to the batch `Loess` type only.
> `StreamingLoess` and `OnlineLoess` remain CPU-only.

---

## Supported Features on GPU

| Feature | CPU | GPU |
| --- | --- | --- |
| Batch fitting | ✅ | ✅ |
| Streaming / Online | ✅ | ❌ |
| All weight / robustness / scaling methods | ✅ | ✅ |
| Confidence / prediction intervals | ✅ | ✅ |
| Cross-validation (k-fold, LOOCV) | ✅ | ✅ |
| Custom per-observation weights | ✅ | ✅ |

---

## Checking GPU Availability

```{r}
library(rfastloess)

gpu_available()
```

---

## Installing GPU Support

```{r}
library(rfastloess)

install_gpu()           # interactive
install_gpu(yes = TRUE) # non-interactive
```

**Restart R after installation.**

---

## Using the GPU Backend

```{r}
library(rfastloess)
set.seed(42)
n <- 100000
x <- seq(0, 2 * pi, length.out = n)
y <- sin(x) + rnorm(n, sd = 0.3)

model <- Loess(
    fraction  = 0.5,
    iterations = 3,
    backend   = "gpu"
)
result <- fit(model, x, y)
```

---

## Performance Guidelines

| Scenario | Recommendation |
| --- | --- |
| n < 10K | CPU (`parallel = TRUE`) |
| n 10K–100K | Try both; GPU wins for larger fractions |
| n > 100K | GPU backend recommended |
| `fraction = 0.5`, n ≥ 50K | GPU outperforms CPU-parallel |

At n = 1M (`fraction = 0.5`, 3 iterations), the GPU backend is approximately **6.6×** faster than CPU-parallel.
