---
title: "Confidence and Prediction Intervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Confidence and Prediction Intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

![Confidence and prediction intervals](../man/figures/intervals_comparison.svg)

> **Note:** Confidence and prediction intervals are available in **Batch**,
> **Streaming**, and **Online** modes.

| Type | Represents | Width | Use |
| --- | --- | --- | --- |
| **Confidence** | Mean curve uncertainty | Narrow | Where is the true trend? |
| **Prediction** | New-point uncertainty | Wide | Where will new data fall? |

---

## Confidence Intervals

Estimate uncertainty in the smoothed curve itself.

```{r intervals_1}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(fraction = 0.5, confidence_intervals = 0.95)
result <- fit(model, x, y)

# Plot with bands
plot(x, y, pch = 16, col = "gray",
    main = "LOESS with 95% Confidence Intervals")
lines(result$x, result$y, col = "blue", lwd = 2)
lines(result$x, result$confidence_lower, col = "blue", lty = 2)
lines(result$x, result$confidence_upper, col = "blue", lty = 2)
```

---

## Prediction Intervals

Wider than confidence intervals — cover where new observations will fall.

```{r intervals_2}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(fraction = 0.5, prediction_intervals = 0.95)
result <- fit(model, x, y)

plot(x, y, pch = 16, col = "gray",
    main = "LOESS with 95% Prediction Intervals")
lines(result$x, result$y, col = "blue", lwd = 2)
lines(result$x, result$prediction_lower, col = "red", lty = 2)
lines(result$x, result$prediction_upper, col = "red", lty = 2)
```

---

## Both Intervals Together

```{r intervals_3}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(
    fraction = 0.5,
    confidence_intervals = 0.95,
    prediction_intervals = 0.95
)
result <- fit(model, x, y)

plot(x, y, pch = 16, col = "gray",
    main = "Confidence vs Prediction Intervals")
lines(result$x, result$y, col = "blue", lwd = 2)

# Confidence interval (narrow, blue)
lines(result$x, result$confidence_lower, col = "blue", lty = 2)
lines(result$x, result$confidence_upper, col = "blue", lty = 2)

# Prediction interval (wide, red)
lines(result$x, result$prediction_lower, col = "red", lty = 3)
lines(result$x, result$prediction_upper, col = "red", lty = 3)

legend("topright",
        c("Data", "Smoothed", "95% CI", "95% PI"),
        pch = c(16, NA, NA, NA), lty = c(NA, 1, 2, 3),
        col = c("gray", "blue", "blue", "red"))
```

---

## Choosing a Coverage Level

```{r intervals_4}
# 99% confidence intervals
model <- Loess(fraction = 0.5, confidence_intervals = 0.99)
result <- fit(model, x, y)
cat("First 6 smoothed values (99% CI):\n")
print(head(result$y))
```

```{r session-info}
sessionInfo()
```
