---
title: "Merge Strategies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Merge Strategies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

Streaming LOESS processes data in fixed-size chunks with a configurable overlap.
Points inside the overlap zone receive estimates from both adjacent chunks; the
merge strategy decides how those two estimates are reconciled.

```text
Chunk A:   [=========|=====]
Chunk B:            [=====|=========]
Overlap:            [=====]
```

| Strategy | Method | Robustness | Speed |
| --- | --- | --- | --- |
| `"average"` | Simple mean of both estimates | Low | Fastest |
| `"take_first"` | Left-chunk estimate only | Low | Fastest |
| `"take_last"` | Right-chunk estimate only | Low | Fastest |
| `"weighted_average"` | Distance-weighted mean | High | Moderate |

![Merge strategy comparison](../man/figures/merge_comparison.svg)

---

## Weighted Average (Default)

Assigns each overlap point a weight proportional to its proximity to the centre
of its respective chunk: points near the left-chunk centre favour the left
estimate; points near the right-chunk centre favour the right. Minimises
boundary artefacts.

**Use when**: Minimising boundary artefacts is more important than speed;
moderate overlap (10–20% of chunk size) is used.

```{r merge_1}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)
x_chunk <- x[seq_len(50)]
y_chunk <- y[seq_len(50)]

model <- StreamingLoess(
    merge_strategy = "weighted_average",
    chunk_size = 5000,
    overlap = 500
)
result <- process_chunk(model, x_chunk, y_chunk)
cat("First 6 smoothed values (weighted_average strategy):\n")
print(head(result$y))
```

---

## Average

Takes the arithmetic mean of the left-chunk and right-chunk estimates in the
overlap region. Fast and sufficient when both chunks have similar smoothing
quality.

**Use when**: Chunks are large and the overlap region has uniform data density.

```{r merge_2}
model <- StreamingLoess(
    merge_strategy = "average",
    chunk_size = 5000,
    overlap = 500
)
result <- process_chunk(model, x_chunk, y_chunk)
cat("First 6 smoothed values (average strategy):\n")
print(head(result$y))
```

---

## Take First

Keeps only the left-chunk estimate in the overlap zone and discards the
right-chunk estimate. Produces a left-flush output.

**Use when**: You need final output values immediately after each chunk (no
look-ahead revision); left-chunk context dominates.

```{r merge_3}
model <- StreamingLoess(
    merge_strategy = "take_first",
    chunk_size = 5000,
    overlap = 500
)
result <- process_chunk(model, x_chunk, y_chunk)
cat("First 6 smoothed values (take_first strategy):\n")
print(head(result$y))
```

---

## Take Last

Keeps only the right-chunk estimate in the overlap zone. The right chunk sees
more of the surrounding data, so its estimates can be more accurate in the
overlap region.

**Use when**: Right-chunk context improves overlap quality; you are
post-processing complete data rather than streaming in real time.

```{r merge_4}
model <- StreamingLoess(
    merge_strategy = "take_last",
    chunk_size = 5000,
    overlap = 500
)
result <- process_chunk(model, x_chunk, y_chunk)
cat("First 6 smoothed values (take_last strategy):\n")
print(head(result$y))
```

---

## Choosing Chunk Size and Overlap

```{r merge_5}
# Rule of thumb: overlap = fraction * chunk_size
fraction  <- 0.3
chunk_size <- 5000
overlap    <- ceiling(fraction * chunk_size)  # 1500

model <- StreamingLoess(
    fraction       = fraction,
    chunk_size     = chunk_size,
    overlap        = overlap,
    merge_strategy = "weighted_average"
)
cat(sprintf("overlap = %d points (%.0f%% of chunk)\n",
            overlap, 100 * overlap / chunk_size))
```

```{r session-info}
sessionInfo()
```
