---
title: "Parameters Reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parameters Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## All Parameters

| Parameter | Default | Range/Options | Description | Mode |
| --- | --- | --- | --- | --- |
| **fraction** | 0.67 | (0, 1] | Smoothing span | All |
| **iterations** | 3 | [0, 1000] | Robustness iterations | All |
| **degree** | 1 | 0–4 | Polynomial degree | All |
| **surface_mode** | `"interpolation"` | 2 options | Fit vs interpolate | All |
| **weight_function** | `"tricube"` | 7 options | Distance kernel | All |
| **robustness_method** | `"bisquare"` | 3 options | Outlier weighting | All |
| **zero_weight_fallback** | `"use_local_mean"` | 3 options | Zero-weight behaviour | All |
| **boundary_policy** | `"extend"` | 4 options | Edge handling | All |
| **scaling_method** | `"mad"` | 3 options | Scale estimation | All |
| **auto_converge** | `NULL` | tolerance | Early stopping | All |
| **parallel** | `TRUE` | `TRUE`/`FALSE` | Multi-threaded execution | Batch |
| **custom_weights** | `NULL` | positive | Per-observation weights | Batch |
| **return_residuals** | `FALSE` | logical | Include residuals | All |
| **return_robustness_weights** | `FALSE` | logical | Include weights | All |
| **return_diagnostics** | `FALSE` | logical | Include metrics | All |
| **confidence_intervals** | `NULL` | (0, 1) | CI level | Batch |
| **prediction_intervals** | `NULL` | (0, 1) | PI level | Batch |
| **distance_metric** | `"normalized"` | string | Distance metric | All |
| **weighted_metric_weights** | `NULL` | numeric vector | Per-dimension distance weights | All |
| **cell** | `NULL` | (0, ∞) | Interpolation cell size | All |
| **interpolation_vertices** | `NULL` | integer | Interpolation grid vertices | All |
| **boundary_degree_fallback** | `FALSE` | logical | Degree fallback at boundaries | All |
| **cv_method** | `NULL` | method | Auto-select fraction | Batch |
| **cv_k** | 5 | [2, ∞) | K-fold count | Batch |
| **cv_fractions** | `NULL` | numeric | Fractions to evaluate | Batch |
| **cv_seed** | `NULL` | integer | CV fold randomisation seed | Batch |
| **chunk_size** | 5000 | [10, ∞) | Points per chunk | Streaming |
| **overlap** | 500 | [0, chunk) | Overlap between chunks | Streaming |
| **merge_strategy** | `"weighted_average"` | 4 options | Merge overlaps | Streaming |
| **window_capacity** | 1000 | [3, ∞) | Max window size | Online |
| **min_points** | 2 | [2, window] | Min before output | Online |
| **update_mode** | `"incremental"` | 2 options | Update strategy | Online |

---

## Parameter Options

| Parameter | Available Options |
| --- | --- |
| **degree** | `0`/`"constant"`, `1`/`"linear"` (default), `2`/`"quadratic"`, `3`/`"cubic"`, `4`/`"quartic"` |
| **surface_mode** | `"interpolation"` (default), `"direct"` |
| **weight_function** | `"tricube"`, `"epanechnikov"`, `"gaussian"`, `"biweight"`, `"cosine"`, `"triangle"`, `"uniform"` |
| **robustness_method** | `"bisquare"`, `"huber"`, `"talwar"` |
| **zero_weight_fallback** | `"use_local_mean"`, `"return_original"`, `"return_none"` |
| **boundary_policy** | `"extend"`, `"reflect"`, `"zero"`, `"noboundary"` |
| **scaling_method** | `"mad"`, `"mar"`, `"mean"` |
| **distance_metric** | `"normalized"`, `"euclidean"`, `"weighted"` |
| **merge_strategy** | `"average"`, `"weighted_average"`, `"take_first"`, `"take_last"` |
| **update_mode** | `"incremental"`, `"full"` |

---

## Core Parameters

### fraction

The proportion of data used for each local fit. **Most important parameter.**

| Value | Effect | Use Case |
| --- | --- | --- |
| 0.1–0.3 | Fine detail | Rapidly changing signals |
| 0.3–0.5 | Balanced | General purpose |
| 0.5–0.7 | Heavy smoothing | Noisy data |
| 0.7–1.0 | Very smooth | Trend extraction |

```{r}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(fraction = 0.3)
result <- fit(model, x, y)
```

### iterations

Number of robustness iterations. Each iteration downweights high-residual points.

```{r}
model <- Loess(iterations = 3L)
result <- fit(model, x, y)
```

### degree

Polynomial degree for local fits. See the [Polynomial Degree](degree.html) vignette.

```{r}
model <- Loess(degree = 2L)  # quadratic
result <- fit(model, x, y)
```

### surface_mode

Controls whether LOESS fits directly at every query point (`"direct"`) or fits at a coarse grid and interpolates (`"interpolation"`, default). Interpolation is faster for large datasets.

```{r}
model <- Loess(surface_mode = "direct")
result <- fit(model, x, y)
```

### distance_metric

Metric used when computing neighbour distances in multivariate mode. Use `"weighted"` with `weighted_metric_weights` to scale each dimension separately.

```{r}
# 2D fit with per-dimension scaling
x2d <- cbind(seq(0, 1, length.out = 100), seq(0, 2, length.out = 100))
model <- Loess(
    distance_metric = "weighted",
    weighted_metric_weights = c(1.0, 0.5)
)
result <- fit(model, x2d, y)
```

### parallel

Enable parallel CPU execution (multiple cores).

```{r}
model <- Loess(parallel = TRUE)
result <- fit(model, x, y)
```

### return_diagnostics

Return goodness-of-fit diagnostics (R², residuals, etc.).

```{r}
model <- Loess(return_diagnostics = TRUE)
result <- fit(model, x, y)
print(result$diagnostics$r_squared)
```

---

## Streaming Parameters

### chunk_size and overlap

```{r}
model <- StreamingLoess(
    chunk_size = 5000L,
    overlap = 500L,
    merge_strategy = "weighted_average"
)
```

---

## Online Parameters

### window_capacity and min_points

```{r}
model <- OnlineLoess(
    window_capacity = 50L,
    min_points = 5L,
    update_mode = "incremental"
)
```
