---
title: "Introduction to rfastloess"
author: "Amir Valizadeh"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to rfastloess}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 7,
    fig.height = 5
)
```

```{r srr-tags, eval = FALSE, echo = FALSE}
#' @srrstats {G1.0} Package introduction with academic context.
#' @srrstats {G1.4} Examples use roxygen2-documented functions.
#' @srrstats {RE1.2} Documents expected numeric vector input format.
#' @srrstats {RE1.4} Documents LOESS assumptions and use cases.
#' @srrstats {RE4.17} Demonstrates print() method for model objects.
#' @srrstats {RE6.0, RE6.2} Demonstrates plot() method with CI visualization.
```

## Overview

`rfastloess` is a high-performance R package for LOESS (Locally Weighted
Scatterplot Smoothing) built on a Rust backend.

LOESS fits a smooth curve through a scatter plot by performing a sequence of
*local polynomial regressions*. For each point $x_i$, the algorithm selects
the nearest `fraction` of observations, assigns distance-based weights (points
farther away receive lower weight), fits a weighted least-squares polynomial,
and reads off the predicted value at $x_i$. Because no global model is
assumed, LOESS adapts naturally to non-linear and non-stationary data.

Optional robustness iterations guard against outliers: after each pass,
residuals are used to recompute per-observation weights so that high-leverage
outliers are progressively downweighted. This makes `iterations > 0` the right
choice whenever the data may contain measurement errors or anomalies.

**Full documentation**: <https://loess.readthedocs.io/>

## Installation

```{r installation, eval=FALSE}
# From R-universe (pre-built binaries, no Rust required)
install.packages("rfastloess", repos = "https://thisisamirv.r-universe.dev")
```

## Quick Start

### Basic Smoothing

`Loess()` accepts all smoothing options at construction time and returns a
reusable model object. Calling `fit(model, x, y)` runs the algorithm and
returns a `LoessResult` list. Key result fields are `$x` (input x values),
`$y` (smoothed fitted values), `$fraction_used`, and `$iterations_used`. The
`fraction` parameter controls the bandwidth: a larger value produces a smoother
curve at the cost of following local features less closely.

```{r basic_example}
library(rfastloess)

# Generate example data
set.seed(42)
x <- seq(0, 10, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

# Initialize model
model <- Loess(fraction = 0.3)
print(model)

# Fit model
result <- fit(model, x, y)
print(result)

# Quick visualization using the S3 plot method
plot(result, main = "Auto-plot of LoessResult")

# For custom plotting, access components directly
plot(x, y, pch = 16, col = "gray", main = "Manual Overlay")
lines(result$x, result$y, col = "red", lwd = 2)
```

### Robust Smoothing with Intervals

Real-world data often contains a small fraction of gross errors. Setting
`iterations > 0` activates robustness re-weighting: after the initial fit,
large residuals identify likely outliers and their contribution is downweighted
for the next iteration. The bisquare robustness function (the default) reduces
the weight of any point whose residual exceeds about six times the median
absolute deviation to zero.

Setting `confidence_intervals = 0.95` attaches pointwise 95% confidence bands
to the smoothed curve. These reflect uncertainty in the mean response, not
individual observations; `prediction_intervals` provides the wider bands
covering future observations.

```{r robust_example}
# Add outliers
y_outliers <- y
y_outliers[sample(1:100, 10)] <- y_outliers[sample(1:100, 10)] + 5

# Robust smoothing with confidence intervals
result_robust <- fit(Loess(
    fraction = 0.3,
    iterations = 5,
    confidence_intervals = 0.95
), x, y_outliers)

# Plot
plot(x, y_outliers, pch = 16, col = "gray", main = "Robust LOESS")
lines(result_robust$x, result_robust$y, col = "red", lwd = 2)
lines(result_robust$x, result_robust$confidence_lower, col = "blue", lty = 2)
lines(result_robust$x, result_robust$confidence_upper, col = "blue", lty = 2)
```

### Diagnostics

When `return_diagnostics = TRUE`, the result includes goodness-of-fit metrics
that are useful for comparing different `fraction` values or kernel choices.
`return_residuals = TRUE` attaches the raw residuals, and
`return_robustness_weights = TRUE` exposes the final per-point robustness
weights so you can see which observations were downweighted.

```{r diagnostics_example}
result_diag <- fit(Loess(
    fraction = 0.3,
    iterations = 3,
    return_diagnostics = TRUE,
    return_residuals = TRUE,
    return_robustness_weights = TRUE
), x, y)

cat("R-squared:  ", result_diag$r_squared, "\n")
cat("RMSE:       ", result_diag$rmse, "\n")
cat("MAE:        ", result_diag$mae, "\n")

# Residual plot
plot(x, result_diag$residuals,
    type = "h", col = "steelblue",
    main = "Residuals", ylab = "Residual"
)
abline(h = 0, lty = 2)
```

## Streaming Processing

`StreamingLoess` processes data in fixed-size chunks and keeps only a small
in-memory buffer, making it practical for datasets that would otherwise
require loading everything into RAM at once. The `chunk_size` parameter
controls how many observations are processed at a time; chunks are stitched
together using a configurable overlap and `merge_strategy` to avoid boundary
artefacts.

After processing all chunks, call `finalize()` to retrieve the combined
`LoessResult` covering the full input range.

```{r streaming_example}
x_full <- seq(0, 10, length.out = 200)
y_full <- sin(x_full) + rnorm(200, sd = 0.2)

model_s <- StreamingLoess(fraction = 0.3, chunk_size = 50L)

# Feed data in chunks of 50
chunk_breaks <- c(0, 50, 100, 150, 200)
for (i in seq_len(length(chunk_breaks) - 1)) {
    idx <- (chunk_breaks[i] + 1):chunk_breaks[i + 1]
    process_chunk(model_s, x_full[idx], y_full[idx])
}

result_s <- finalize(model_s)

plot(x_full, y_full, pch = 16, col = "gray", main = "Streaming LOESS")
lines(result_s$x, result_s$y, col = "red", lwd = 2)
```

## Online Processing

`OnlineLoess` is designed for real-time data streams where observations
arrive one at a time. The model maintains a sliding window of the most recent
`window_capacity` points. `add_point(model, x, y)` adds a single observation
and immediately returns the smoothed value at that point — or `NULL` if fewer
than `min_points` have been seen yet.

The `update_mode` parameter controls how the window is updated:
`"full"` (default) re-smooths all window points after each addition, giving
the most accurate estimate; `"incremental"` updates only the newest point,
which is faster but less accurate.

```{r online_example}
model_o <- OnlineLoess(
    fraction = 0.5,
    window_capacity = 30L,
    min_points = 3L
)

x_stream <- seq(0, 10, length.out = 60)
y_stream <- sin(x_stream) + rnorm(60, sd = 0.2)

smoothed_values <- numeric(length(x_stream))
for (i in seq_along(x_stream)) {
    out <- add_point(model_o, x_stream[i], y_stream[i])
    smoothed_values[i] <- if (is.null(out)) NA_real_ else out$smoothed
}

plot(x_stream, y_stream, pch = 16, col = "gray", main = "Online LOESS")
lines(x_stream[!is.na(smoothed_values)],
    smoothed_values[!is.na(smoothed_values)],
    col = "red", lwd = 2
)
```

## Main Classes

| Class            | Use Case                                     |
|------------------|----------------------------------------------|
| `Loess`          | Primary interface — batch processing         |
| `StreamingLoess` | Large datasets processed chunk by chunk      |
| `OnlineLoess`    | Real-time streams, one observation at a time |

## Learn More

For comprehensive documentation including:

- Parameter selection guides
- Streaming and online processing tutorials
- Genomic data examples
- Performance benchmarks

Visit: **<https://loess.readthedocs.io/>**

## Session Info

```{r session_info}
sessionInfo()
```
