---
title: "Scaling Methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scaling Methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

When `iterations > 0`, LOESS computes robustness weights by comparing each
residual to the current residual scale estimate $\hat{\sigma}$. The
`scaling_method` parameter controls how $\hat{\sigma}$ is estimated.

$$w_i = B\!\left(\frac{|r_i|}{6 \cdot \hat{\sigma}}\right)$$

| Method | Formula | Robustness | Speed |
| --- | --- | --- | --- |
| `"mad"` | Median absolute deviation from median | Very robust | Moderate |
| `"mar"` | Median of \|residuals\| | Robust | Fast |
| `"mean"` | Mean of \|residuals\| | Less robust | Fastest |

![Scaling method comparison](../man/figures/scaling_comparison.svg)

---

## MAD — Median Absolute Deviation (Default)

$$\hat{\sigma} = \text{median}(|r_i - \text{median}(r_i)|)$$

First centers residuals at their median, then takes the median of the absolute
deviations. Double use of the median makes it highly resistant to extreme
outliers.

**Use when**: Data may contain outliers (default for most applications).

```{r scaling_1}
library(rfastloess)
set.seed(42)
x <- seq(0, 2 * pi, length.out = 100)
y <- sin(x) + rnorm(100, sd = 0.3)

model <- Loess(iterations = 3, scaling_method = "mad")
result <- fit(model, x, y)
print(head(result$y))
```

---

## MAR — Median Absolute Residual

$$\hat{\sigma} = \text{median}(|r_i|)$$

Uses the uncentered median — unlike MAD it does not subtract the residual median
first. Still robust (median-based), faster (one partial sort instead of two).

**Use when**: Speed matters and data have minimal systematic bias in residuals.

```{r scaling_2}
model <- Loess(iterations = 3, scaling_method = "mar")
result <- fit(model, x, y)
print(head(result$y))
```

---

## Mean — Mean Absolute Residual

$$\hat{\sigma} = \frac{1}{n}\sum_i |r_i|$$

Arithmetic mean of absolute residuals. Non-robust: a single extreme outlier
inflates $\hat{\sigma}$, causing the bisquare weight function to
under-downweight other outliers.

**Use when**: Clean data with no outliers; maximum computation speed required.

```{r scaling_3}
model <- Loess(iterations = 3, scaling_method = "mean")
result <- fit(model, x, y)
print(head(result$y))
```

---

## Comparing Scaling Methods

```{r scaling_4}
library(rfastloess)
set.seed(42)
x <- 1:100
y <- sin(x / 10) + rnorm(100, sd = 0.3)
y[c(20, 50, 80)] <- y[c(20, 50, 80)] + 5  # outliers

methods <- c("mad", "mar", "mean")
colors  <- c("blue", "red", "green")

plot(x, y, pch = 16, col = "gray",
     main = "Scaling Method Comparison (with outliers)")

for (i in seq_along(methods)) {
    model  <- Loess(iterations = 3, scaling_method = methods[i])
    result <- fit(model, x, y)
    lines(result$x, result$y, col = colors[i], lwd = 2)
}

legend("topright", methods, col = colors, lwd = 2)
```

```{r session-info}
sessionInfo()
```
