---
title: "Use Case: Real-Time Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use Case: Real-Time Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- markdownlint-disable MD024 -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

When data arrives continuously — from sensors, logs, or streaming pipelines —
you need incremental smoothing that doesn't require reprocessing the entire
dataset.

---

## Online Mode: Point-by-Point

For true real-time applications where each point must be processed
immediately.

`window_capacity = 25` limits the internal buffer to the 25 most recent
observations; each `add_point` call costs O(window) rather than growing with
total history. `min_points = 5` suppresses output until the window holds enough
points for a stable fit — calls made before that threshold return `NULL`.
`update_mode = "incremental"` re-fits only the most recent point rather than the
full window, halving typical latency at a modest accuracy cost.

### Sensor Data Example

```{r use_case_real_time_1}
library(rfastloess)

set.seed(42)
times        <- 1:100
temperatures <- 20 + 5 * sin(times / 10) + rnorm(100)

model <- OnlineLoess(
    fraction = 0.3,
    window_capacity = 25,
    min_points = 5,
    update_mode = "incremental"
)
for (i in seq_along(times)) {
    result <- add_point(model, times[i], temperatures[i])
    if (!is.null(result))
        cat(sprintf("Time %d: %.2f\n", times[i], result$y))
    if (i >= 10) break  # print only the first few outputs
}
```

### Accumulating Results

```{r use_case_real_time_2}
library(rfastloess)
set.seed(42)
n      <- 200
times  <- seq_len(n)
signal <- 10 + 2 * sin(times / 20) + rnorm(n, sd = 1)

model <- OnlineLoess(
    fraction        = 0.3,
    window_capacity = 30,
    min_points      = 5
)

smoothed_x <- numeric(n)
smoothed_y <- numeric(n)
n_out <- 0L

for (i in seq_len(n)) {
    result <- add_point(model, times[i], signal[i])
    if (!is.null(result)) {
        n_out <- n_out + 1L
        smoothed_x[n_out] <- times[i]
        smoothed_y[n_out] <- result$y
    }
}

smoothed_x <- smoothed_x[seq_len(n_out)]
smoothed_y <- smoothed_y[seq_len(n_out)]

plot(times, signal, pch = 16, cex = 0.4, col = "gray",
    xlab = "Time", ylab = "Value",
    main = "Online LOESS — Accumulated Output")
lines(smoothed_x, smoothed_y, col = "blue", lwd = 2)
```

---

## Streaming Mode: Chunk Processing

For large datasets that arrive in batches or files.

`chunk_size` controls how many data points are processed in one pass;
matching it to your file-read buffer or message-batch size avoids
unnecessary copying. `overlap` retains that many points from the previous
chunk as context so the neighbourhood at chunk boundaries is not
artificially truncated. `merge_strategy = "weighted_average"` blends the
overlapping region smoothly; use `"take_last"` if chunk boundaries are
guaranteed to be well separated and no blending is needed.

> **Always call finalize():** The streaming adapter buffers overlap data.
> Call `finalize(model)` after the last chunk to retrieve the buffered tail.

### Log File Processing

```{r use_case_real_time_3}
library(rfastloess)
set.seed(42)

total_points <- 100000
x <- 0:(total_points - 1)
y <- sin(x / 1000) + rnorm(total_points, sd = 0.1)

model <- StreamingLoess(
    fraction       = 0.05,
    chunk_size     = 10000,
    overlap        = 1000,
    merge_strategy = "weighted_average"
)
process_chunk(model, x, y)
result <- finalize(model)
cat("Processed", length(result$y), "points\n")
```

---

## Real-Time Dashboard Example

The dashboard pattern uses a plain LOESS fit on a manually managed sliding
window rather than `OnlineLoess`. This is the simplest approach when your UI
framework already owns the data buffer and you only need the most recent
smoothed value per frame. The trade-off is a full O(window^2) refit on every
tick; for high-frequency streams prefer `OnlineLoess` with `update_mode =
"incremental"` to bound per-frame cost.

```{r use_case_real_time_4}
library(rfastloess)
set.seed(42)

window_capacity <- 50
data_x <- numeric(0)
data_y <- numeric(0)

for (i in 0:199) {
    xi <- i
    yi <- 25.0 + 10 * sin(i / 20) + rnorm(1, sd = 2)
    data_x <- c(data_x, xi)
    data_y <- c(data_y, yi)

    if (length(data_x) > window_capacity) {
        data_x <- tail(data_x, window_capacity)
        data_y <- tail(data_y, window_capacity)
    }

    if (length(data_x) >= 5) {
        model <- Loess(fraction = 0.4)
        result <- fit(model, data_x, data_y)
        current_smoothed <- tail(result$y, 1)
    }
}
```

---

## Choosing Parameters

### Online Mode

| Parameter | Guidance |
| --- | --- |
| `window_capacity` | Enough history for `fraction` to work |
| `min_points` | 2–5 typically; higher for stability |
| `update_mode` | `"incremental"` for speed, `"full"` for accuracy |

### Streaming Mode

| Parameter | Guidance |
| --- | --- |
| `chunk_size` | Balance memory vs. processing overhead |
| `overlap` | 10–20% of chunk_size for smooth transitions |
| `merge_strategy` | `"weighted_average"` (quality) vs `"average"` (simple) |

---

## Performance Considerations

| Mode | Memory | Latency | Use Case |
| --- | --- | --- | --- |
| **Online** | Fixed (window) | ~1ms/point | Sensors, dashboards |
| **Streaming** | ~chunk_size | ~100ms/chunk | Large files, ETL |
| **Batch** | Full dataset | N/A | Analysis, reports |

---

## See Also

- `vignette("adapter-choice", package = "rfastloess")` — Detailed mode
  comparison
- `vignette("merge", package = "rfastloess")` — Chunk reconciliation in depth
- `vignette("scaling", package = "rfastloess")` — Robustness scale estimation
- `vignette("use-case-time-series", package = "rfastloess")` — General
  time series analysis

```{r session-info}
sessionInfo()
```
