---
title: "Use Case: Time Series Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use Case: Time Series Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- markdownlint-disable MD024 -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

LOESS provides flexible trend extraction from time series without parametric
assumptions.

---

## Basic Trend Extraction

`fraction = 0.1` sizes the neighbourhood as 10% of the data at each evaluation
point — narrow enough to follow a slowly varying trend without smearing periodic
variation. Three robustness `iterations` down-weight noise spikes so they cannot
bias the fitted curve; this is especially important when the signal-to-noise
ratio is low or when occasional outliers are expected.

```{r use_case_time_series_1}
library(rfastloess)

set.seed(42)
t <- seq(0, 100, length.out = 500)
trend <- 10 + 0.5 * t + 3 * sin(t / 10)
noise <- rnorm(500, sd = 3)
y <- trend + noise

model <- Loess(fraction = 0.1, iterations = 3)
result <- fit(model, t, y)

plot(t, y, col = "gray", pch = ".",
    xlab = "Time", ylab = "Value", main = "Trend Extraction")
lines(result$x, result$y, col = "blue", lwd = 2)
legend("topleft", c("Observed", "Trend (LOESS)"),
        pch = c(1, NA), lty = c(NA, 1), col = c("gray", "blue"))
```

---

## Seasonal Decomposition

Setting `return_residuals = TRUE` stores `observed − smoothed` alongside the
smooth. A slightly wider `fraction = 0.4` produces a smoother baseline trend, so
short-duration oscillations end up in the residuals rather than being absorbed
into the trend component. The residual series is then ready for spectral
analysis, seasonality detection, or change-point methods.

```{r use_case_time_series_2}
library(rfastloess)
set.seed(42)

t        <- 1:120
trend    <- 100 + 0.5 * t
seasonal <- 15 * sin(2 * pi * t / 12)
noise    <- rnorm(120, sd = 5)
y        <- trend + seasonal + noise

model    <- Loess(fraction = 0.4, iterations = 2)
result   <- fit(model, t, y)
residual <- y - result$y

par(mfrow = c(2, 1), mar = c(4, 4, 2, 0.5))
plot(t, y, type = "l", col = "gray",
    main = "Original + LOESS Trend", xlab = "Month", ylab = "Value")
lines(result$x, result$y, col = "blue", lwd = 2)

plot(t, residual, type = "l", col = "darkgreen",
    main = "Detrended Residual (Seasonal Component)",
    xlab = "Month", ylab = "Residual")
abline(h = 0, lty = 2, col = "red")
par(mfrow = c(1, 1))
```

---

## Irregular Time Grids

```{r use_case_time_series_3}
library(rfastloess)
set.seed(42)

t_irregular <- c(sort(runif(200, 0, 50)), sort(runif(50, 50, 100)))
y_irregular <- sin(t_irregular / 10) + rnorm(length(t_irregular), sd = 0.5)

model  <- Loess(fraction = 0.2, iterations = 2)
result <- fit(model, t_irregular, y_irregular)

plot(t_irregular, y_irregular, pch = 16, cex = 0.4, col = "gray",
    xlab = "Time", ylab = "Value", main = "Irregularly Sampled Time Series")
lines(result$x, result$y, col = "blue", lwd = 2)
```

---

## Forecasting with Prediction Intervals

Prediction intervals widen the uncertainty band to include both the
uncertainty in the fitted curve (confidence interval) and the expected
scatter of new observations around it. `fraction = 0.2` offers a balance
between local detail and stable interval width — too small a fraction
produces jagged interval edges; too large a fraction underestimates local
variance near turning points.

```{r use_case_time_series_4}
library(rfastloess)
set.seed(42)
t <- seq(0, 100, length.out = 500)
y <- 10 + 0.3 * t + sin(t / 5) + rnorm(500, sd = 2)

model <- Loess(
    fraction = 0.2,
    iterations = 3,
    confidence_intervals = 0.95,
    prediction_intervals = 0.95
)
result <- fit(model, t, y)

plot(t, y, pch = ".", col = "gray",
    xlab = "Time", ylab = "Value",
    main = "Trend with 95% Prediction Interval")
polygon(c(result$x, rev(result$x)),
        c(result$prediction_upper, rev(result$prediction_lower)),
        col = rgb(0, 0, 1, 0.15), border = NA)
lines(result$x, result$y, col = "blue", lwd = 2)
```

---

## Handling Missing Data

LOESS naturally handles irregular time sampling:

```{r use_case_time_series_5}
library(rfastloess)
set.seed(42)

t_irregular <- sort(runif(200, 0, 100))
y_irregular <- 10 + t_irregular * 0.3 + rnorm(200, sd = 2)

model <- Loess(fraction = 0.2)
result <- fit(model, t_irregular, y_irregular)
cat("Smoothed y[0]:", result$y[1], "\n")
```

---

## Multi-Scale Analysis

Use different fractions to extract features at different scales:

```{r use_case_time_series_6}
library(rfastloess)
set.seed(42)
t <- seq(0, 100, length.out = 500)
trend_true <- 10 + 0.5 * t + 3 * sin(t / 10)
y <- trend_true + rnorm(length(t), sd = 3)

fractions <- c(0.05, 0.2, 0.5)

plot(t, y, col = "gray", pch = ".",
    xlab = "Time", ylab = "Value", main = "Multi-Scale LOESS")

for (f in fractions) {
    model <- Loess(fraction = f)
    result <- fit(model, t, y)
    lines(result$x, result$y, lwd = 2)
}
legend("topleft", paste0("fraction=", fractions), lty = 1)
```

---

## Gene Expression Time Course

Biological application:

```{r use_case_time_series_7}
library(rfastloess)
set.seed(42)

hours <- seq(0, 24, by = 0.5)
expression <- 100 * (1 + 0.5 * sin(hours * pi / 12)) +
    rnorm(length(hours), sd = 10)

model <- Loess(
    fraction = 0.3,
    iterations = 3,
    confidence_intervals = 0.95,
    return_diagnostics = TRUE
)
result <- fit(model, hours, expression)

cat("R2:", result$diagnostics$r_squared, "\n")
```

---

## Choosing Fraction for Time Series

| Data Type | Recommended Fraction | Rationale |
| --- | --- | --- |
| Daily data (years) | 0.3–0.5 | Capture annual trends |
| Hourly data (days) | 0.1–0.2 | Capture daily patterns |
| Sensor data (minutes) | 0.05–0.1 | Preserve short-term features |
| Noisy data | Higher | Reduce noise impact |
| Clean data | Lower | Preserve detail |

---

## See Also

- `vignette("use-case-real-time", package = "rfastloess")` — For
  streaming time series
- `vignette("cross-validation", package = "rfastloess")` — Optimal
  fraction selection
- `vignette("degree", package = "rfastloess")` — Degree 2 for curved trends
- `vignette("boundary", package = "rfastloess")` — Edge bias in trend extraction
- `?Loess` — Full parameter reference

```{r session-info}
sessionInfo()
```
