---
title: "Use Case: Real-Time Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use Case: Real-Time Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Overview

The Online adapter provides incremental smoothing for continuously arriving data.

---

## Online Mode: Point-by-Point

`window_capacity = 25` limits the internal buffer to the 25 most recent observations; each `add_point` call costs O(window) rather than growing with total history. `min_points = 5` suppresses output until the window holds enough points for a stable fit — calls made before that threshold return `NULL`. `update_mode = "incremental"` re-fits only the most recent point rather than the full window, halving typical latency at a modest accuracy cost.

### Sensor Data Example

```{r use_case_real_time_1}
library(rfastloess)

set.seed(42)
times        <- 1:100
temperatures <- 20 + 5 * sin(times / 10) + rnorm(100)

model <- OnlineLoess(
    fraction = 0.3,
    window_capacity = 25,
    min_points = 5,
    update_mode = "incremental"
)
for (i in seq_along(times)) {
    result <- add_point(model, times[i], temperatures[i])
    if (!is.null(result))
        cat(sprintf("Time %d: %.2f\n", times[i], result$y))
}
```

---

## Accumulating Results

```{r use_case_real_time_2}
library(rfastloess)
set.seed(42)
n      <- 200
times  <- seq_len(n)
signal <- 10 + 2 * sin(times / 20) + rnorm(n, sd = 1)

model <- OnlineLoess(
    fraction        = 0.3,
    window_capacity = 30,
    min_points      = 5
)

smoothed_x <- numeric(n)
smoothed_y <- numeric(n)
n_out <- 0L

for (i in seq_len(n)) {
    result <- add_point(model, times[i], signal[i])
    if (!is.null(result)) {
        n_out <- n_out + 1L
        smoothed_x[n_out] <- times[i]
        smoothed_y[n_out] <- result$y
    }
}

smoothed_x <- smoothed_x[seq_len(n_out)]
smoothed_y <- smoothed_y[seq_len(n_out)]

plot(times, signal, pch = 16, cex = 0.4, col = "gray",
     xlab = "Time", ylab = "Value",
     main = "Online LOESS — Accumulated Output")
lines(smoothed_x, smoothed_y, col = "blue", lwd = 2)
```

---

## Update Modes

| Mode | Behaviour | Latency | Accuracy |
| --- | --- | --- | --- |
| `"incremental"` | Re-fits only the newest point | Low | Moderate |
| `"full"` | Re-fits the entire window | Higher | Higher |

---

## Streaming Mode for Large Batches

```{r use_case_real_time_3}
library(rfastloess)

model <- StreamingLoess(
    fraction       = 0.3,
    iterations     = 2,
    chunk_size     = 5000,
    overlap        = 500,
    merge_strategy = "weighted_average"
)

set.seed(42)
n_total <- 20000
x_all   <- seq_len(n_total)
y_all   <- sin(x_all / 500) + rnorm(n_total, sd = 0.5)

chunk_size <- 5000
n_chunks   <- ceiling(n_total / chunk_size)

for (i in seq_len(n_chunks)) {
    idx_from <- (i - 1) * chunk_size + 1
    idx_to   <- min(i * chunk_size, n_total)
    result   <- process_chunk(model, x_all[idx_from:idx_to],
                                      y_all[idx_from:idx_to])
}

final <- finalize(model)
```
